---
title: "Simulation experiments for iCorShrink2 - Hub"
author: "Kushal K Dey, Rahul Mazumdar"
date: "1/14/2019"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(MASS) 
library(Matrix)
library(corrplot)
library(CorShrink)
library(corpcor)
library(scales)
library(gridExtra)
library(ggplot2)
library(CVXR)
```


We simulate data from a Hub correlation matrix

## Simulation - Hub correlation

```{r}
P <- 50
block <- 5
mat <- 0.3*diag(1,block) + 0.7*rep(1,block) %*% t(rep(1, block))
Sigma <-   bdiag(mat, mat, mat, mat, mat, mat, mat, mat, mat, mat)
corSigma <- cov2cor(Sigma)
pcorSigma <- as.matrix(cor2pcor(corSigma))
```

Correlation plot

```{r}
col2 <- c("blue", "white", "red")
corrplot(as.matrix(corSigma), diag = FALSE, col = colorRampPalette(col2)(200), tl.pos = "td", 
         tl.col = "black", tl.cex = 0.8, rect.col = "white", 
         na.label.col = "white", method = "color", type = "upper")
```

Partial correlation plot

```{r}
col2 <- c("blue", "white", "red")
corrplot(pcorSigma, diag = FALSE, col = colorRampPalette(col2)(200), tl.pos = "td", 
         tl.col = "black", tl.cex = 0.8, rect.col = "white", 
         na.label.col = "white", method = "color", type = "upper")
```

Note that for positively correlated blocks or hub structure in the correlations, both the correlation and the partial correlation matrix show hub patterns, the latter being weaker.

##  Generating data and adding missing entries

We incorporate 50 % missing entries

```{r}
data = mvrnorm(n=500, rep(0, P), Sigma)
data_with_missing = apply(data, c(1, 2), function(x) {
        if(runif(1,0,1) < 0.5){
          y = NA
        }else{
          y = x
        }
        return(y)
})
```

## Pairwise sample correlation matrix

```{r}
cor2 = cor(data_with_missing, use = "pairwise.complete.obs")
col2 <- c("blue", "white", "red")
corrplot(as.matrix(cor2), diag = FALSE,
         col = colorRampPalette(col2)(200),
         tl.pos = "td", tl.cex = 0.4, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "upper")
```

## iCorSrhink estimate

### Expo = 1

```{r}
alphas <- c(0.0001, 0.001, 0.01, 0.1, 1, 10)
Partial.est <- lapply(alphas,
                function(alpha) {
                  out = iCorShrink2Data(data_with_missing, alpha = alpha)
                })
frob_dist = c()
for (m in 1:6){
  frob_dist[m] = mean((Partial.est[[m]] - pcorSigma)^2)
}
```

```{r}
par(mfrow=c(3,2))
for(m in 1:6){
  corrplot(as.matrix(cov2cor(Partial.est[[m]])), diag = FALSE,
         col = colorRampPalette(col2)(200),
         tl.pos = "td", tl.cex = 0.4, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "upper")
}
```

```{r}
frob_dist
```


### alpha = 0.1 different exponents

```{r}
expos = c(1e-05, 0.001, 0.05, 1, 2, 10)
alpha = 0.1
Partial.est <- lapply(expos,
                function(expo) {
                  out = iCorShrink2Data(data_with_missing, alpha = alpha, expo = expo)
                })
frob_dist = c()
for (m in 1:6){
  frob_dist[m] = mean((Partial.est[[m]] - pcorSigma)^2)
}
min(frob_dist[!is.na(frob_dist)])
par(mfrow=c(3,2))
for(m in 1:6){
  corrplot(as.matrix(cov2cor(Partial.est[[m]])), diag = FALSE,
         col = colorRampPalette(col2)(200),
         tl.pos = "td", tl.cex = 0.4, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "upper")
}
```

```{r}
frob_dist
```

