---
title: "Simulation experiments for iCorShrink"
author: "Kushal K Dey, Rahul Mazumdar"
date: "12/19/2018"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(MASS) 
library(Matrix)
library(corrplot)
library(CorShrink)
library(corpcor)
library(scales)
library(gridExtra)
library(ggplot2)
library(CVXR)
```

We simulate data from a Hub correlation matrix

## Simulation - Hub correlation

```{r}
P <- 50
block <- 5
mat <- 0.3*diag(1,block) + 0.7*rep(1,block) %*% t(rep(1, block))
Sigma <-   bdiag(mat, mat, mat, mat, mat, mat, mat, mat, mat, mat)
corSigma <- cov2cor(Sigma)
pcorSigma <- as.matrix(cor2pcor(corSigma))
```

Correlation plot

```{r}
col2 <- c("blue", "white", "red")
corrplot(as.matrix(corSigma), diag = FALSE, col = colorRampPalette(col2)(200), tl.pos = "td", 
         tl.col = "black", tl.cex = 0.8, rect.col = "white", 
         na.label.col = "white", method = "color", type = "upper")
```

Partial correlation plot

```{r}
col2 <- c("blue", "white", "red")
corrplot(pcorSigma, diag = FALSE, col = colorRampPalette(col2)(200), tl.pos = "td", 
         tl.col = "black", tl.cex = 0.8, rect.col = "white", 
         na.label.col = "white", method = "color", type = "upper")
```

Note that for positively correlated blocks or hub structure in the correlations, both the correlation and the partial correlation matrix show hub patterns, the latter being weaker.

##  Generating data and adding missing entries

We incorporate 50 % missing entries

```{r}
data = mvrnorm(n=500, rep(0, P), Sigma)
data_with_missing = apply(data, c(1, 2), function(x) {
        if(runif(1,0,1) < 0.5){
          y = NA
        }else{
          y = x
        }
        return(y)
})
```

## Pairwise sample correlation matrix

```{r}
cor2 = cor(data_with_missing, use = "pairwise.complete.obs")
col2 <- c("blue", "white", "red")
corrplot(as.matrix(cor2), diag = FALSE,
         col = colorRampPalette(col2)(200),
         tl.pos = "td", tl.cex = 0.4, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "upper")
```


## iCorSrhink estimate

```{r}
alphas <- c(0.0001, 0.001, 0.01, 0.1, 1, 10)
Partial.est <- lapply(alphas,
                function(alpha) {
                  out = iCorShrinkData(data_with_missing, alpha = alpha)
                })
frob_dist = c()
for (m in 1:6){
  frob_dist[m] = mean((Partial.est[[m]] - pcorSigma)^2)
}
min(frob_dist[!is.na(frob_dist)])
Phat = as.matrix(cov2cor(Partial.est[[which.min(frob_dist)]]))
```


```{r}
corrplot(Phat, diag = FALSE,
         col = colorRampPalette(col2)(200),
         tl.pos = "td", tl.cex = 0.4, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "upper")
```


We find the performance of the estimated matrices to be visually good, with the best performing result corresponds to be $\alpha = 0.1$.

Other methods cannot handle missing data like we do. However, if we use the pairwise covariance matrix and GLASSO, we get the following result.

## One alpha different weights

```{r}
omegas = c(0.3, 0.5, 0.7, 1, 5, 10)
alpha = 0.1
Partial.est <- lapply(omegas,
                function(omega) {
                  out = CorShrink::iCorShrinkData(data_with_missing, alpha = 0.1, lambda = omega)
                })

col2 <- c("blue", "white", "red")
par(mfrow=c(3,2))
for(m in 1:6){
  corrplot(as.matrix(cov2cor(Partial.est[[m]])), diag = FALSE,
         col = colorRampPalette(col2)(200),
         tl.pos = "td", tl.cex = 0.4, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "upper")
}

```


## Comparison with GLASSO

```{r echo=FALSE, eval=TRUE}

binary_indicator = matrix(1, nrow(data_with_missing), ncol(data_with_missing))
binary_indicator[is.na(data_with_missing)]= 0
common_samples = t(binary_indicator)%*%binary_indicator


###############  compute pairwise covariance/correlation matrix   #######################

pairwise_cov = cov(data_with_missing, use = "pairwise.complete.obs")
sigma_vals = sqrt(diag(pairwise_cov))
pairwise_cor = cov2cor(pairwise_cov)
pairwise_cor[is.na(pairwise_cor)] = 0
pairwise_cor[pairwise_cor > 0.95] = 0.95
pairwise_cor[pairwise_cor < -0.95] = -0.95
diag(pairwise_cor) = 1
common_samples[common_samples <= 2] = 2
pairwise_cov = diag(sigma_vals) %*% pairwise_cor %*% diag(sigma_vals)

  
Omega <- Semidef(dim(common_samples)[1])
obj = Minimize(-log_det(Omega) + matrix_trace(Omega %*% pairwise_cov))
Omega.est <- lapply(alphas,
                function(alpha) {
                    constraints <-list(sum(abs(Omega)) <= alpha)
                    ## Form and solve optimization problem
                    prob <- Problem(obj, constraints)
                    result <- solve(prob)
                    
                    ## Create covariance matrix
                    R_hat <- base::solve(result$getValue(Omega))
                    Omega_hat <- result$getValue(Omega)
                    Omega_hat[abs(Omega_hat) <= 1e-4] <- 0
                    Omega_hat
                })

Partial.est = lapply(Omega.est, function(M) {
                              temp = -cov2cor(as.matrix(M))
                              diag(temp) = 1
                              return(temp)}
                    )
frob_dist = c()
for (m in 1:6){
  frob_dist[m] = mean((Partial.est[[m]] - pcorSigma)^2)
}
min(frob_dist[!is.na(frob_dist)])
```

```{r eval=FALSE}
par(mfrow=c(3,2))
for(m in 1:6){
  mat2 = cov2cor(as.matrix(Partial.est[[m]]))
  mat2[mat2 < -1] = -1
  mat2[mat2 > 1] = 1
  corrplot(mat2, diag = FALSE,
         col = colorRampPalette(col2)(200),
         tl.pos = "td", tl.cex = 0.4, tl.col = "black",
         rect.col = "white",na.label.col = "white",
         method = "color", type = "upper")
}

```


Looks like GLASSO is performing worse than the Robustified GLASSO version we propose.

