---
title: "Predictive Accuracy of CorShrink-sparse and CorShrink-Loss approaches on PLIN1 gene expression"
author: "Kushal K Dey, Rahul Mazumdar"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(corrplot)
library(CorShrink)
library(CVXR)
```

## PLIN1 gene data

```{r}
data("sample_by_feature_data")
data("common_samples")
```

```{r}
tau_range = c(seq(0.1, 1, length.out = 10))
measure = c()
train_sample_id = sample(nrow(sample_by_feature_data), 300, replace = FALSE)
predict_sample_id = setdiff(1:nrow(sample_by_feature_data), train_sample_id)

################## Correlation matrix for training data   #########################

train_datamat = sample_by_feature_data[train_sample_id,]
empirical_cor = cor(train_datamat, method = "pearson", use = "pairwise.complete.obs")
empirical_cor[is.na(empirical_cor)] = 0

################# Correlation matrix for prediction data  #########################

predict_datamat = sample_by_feature_data[predict_sample_id,]
cormat2 = cor(predict_datamat, method = "pearson", use = "pairwise.complete.obs")
cormat2[is.na(cormat2)] = 0
```

## CorShrink-sparse and CorShrink on PLIN1

```{r}
out1 = CorShrinkData(train_datamat, sd_boot = FALSE, image = "null",
                               image.control = list(tl.cex = 0.2))
estcor1 = out1$cor

out2 = CorShrink2Data(train_datamat)

estcor2 = out2

```

The Frobenius distance between the estimated correlation matrices from the training data and that estimated from the left out prediction data.

```{r}
mean((estcor1 - cormat2)^2)  #######  CorShrink
mean((estcor2 - cormat2)^2)  #######  CorShrink2

```

## PLIN1 gene correlation visualization

```{r}
data("sample_by_feature_data")
data("common_samples")

list_cors = list()
temp = cor(sample_by_feature_data, use = "pairwise.complete.obs")
colnames(temp)=NULL
rownames(temp) =NULL
list_cors[[1]] = temp
corshrink_out = CorShrinkData(sample_by_feature_data, sd_boot = FALSE, image = "null",
                                image.control = list(tl.cex = 0.2))
temp2 = corshrink_out$cor
colnames(temp2)=NULL
rownames(temp2) =NULL
list_cors[[2]] = temp2
list_cors[[3]] = CorShrink2Data(sample_by_feature_data)
list_cors[[4]] = CorShrink2DataLoss(sample_by_feature_data, alpha = 1e-05)
list_cors[[5]] = CorShrink2DataLoss(sample_by_feature_data, alpha = 1e-02)
list_cors[[6]] = CorShrink2DataLoss(sample_by_feature_data, alpha = 1)
list_cors[[7]] = CorShrink2DataLoss(sample_by_feature_data, alpha = 10)
list_cors[[8]] = CorShrink2DataLoss(sample_by_feature_data, alpha = 100)

par(mfrow = c(4, 2))
for(m in 1:8){
  col2 <- c("blue", "white", "red")
  corrplot(as.matrix(list_cors[[m]]), diag = FALSE, col = colorRampPalette(col2)(200), tl.pos = "td",
         tl.col = "black", tl.cex = 0.8, rect.col = "white",
         na.label.col = "white", method = "color", type = "upper")
}

```


## Multiple genes analysis

Choose a random set of 100 genes from GTEx.

```{r}
person_tissue_genes = get(load("data/person_tissue_genes_voom.rda")) ##  544 x 53 x 16069 '
data("common_samples")
sets_of_genes = sample(dim(person_tissue_genes)[3], 100, replace = FALSE) 

```

Choose a training dataset randomly of 300 individuals and then use that to predict the correlation structure for the left-out individuals.

```{r echo=TRUE, eval=FALSE}
measure = c()
for(numgene in 1:length(sets_of_genes)){
  train_sample_id = sample(nrow(person_tissue_genes), 300, replace = FALSE)
  predict_sample_id = setdiff(1:nrow(person_tissue_genes), train_sample_id)
  train_datamat = person_tissue_genes[train_sample_id,,sets_of_genes[numgene]]
  empirical_cor = cor(train_datamat, method = "pearson", use = "pairwise.complete.obs")
  empirical_cor[is.na(empirical_cor)] = 0
  
  corshrink_out = CorShrinkData(train_datamat, sd_boot = FALSE, image = "null",
                                image.control = list(tl.cex = 0.2))
  corshrink_cor = corshrink_out$cor
  corshrink1_cor = CorShrink2Data(train_datamat)
  corshrink2_cor = CorShrink2DataLoss(train_datamat, alpha = 1e-05)
  corshrink3_cor = CorShrink2DataLoss(train_datamat, alpha = 1e-02)
  corshrink4_cor = CorShrink2DataLoss(train_datamat, alpha = 1)
  corshrink5_cor = CorShrink2DataLoss(train_datamat, alpha = 10)
  

  predict_datamat = person_tissue_genes[predict_sample_id,,sets_of_genes[numgene]]
  cormat2 = cor(predict_datamat, method = "pearson", use = "pairwise.complete.obs")
  cormat2[is.na(cormat2)] = 0
  measure = rbind(measure, c(mean((cormat2 - empirical_cor)^2),
                             mean((cormat2 - corshrink_cor)^2),
                             mean((cormat2 - corshrink1_cor)^2),
                             mean((cormat2 - corshrink2_cor)^2),
                             mean((cormat2 - corshrink3_cor)^2),
                             mean((cormat2 - corshrink4_cor)^2),
                             mean((cormat2 - corshrink5_cor)^2)))
  cat("We are at numgene", numgene)
}

```

```{r}
measdure = get(load("output/PLIN1_predictive_perform_corshrink_sparse_loss.rda"))
```


```{r}
df1 = data.frame("frobenius" = c(measure[,1], measure[,2], measure[,3],
                                 measure[,4], measure[,5], measure[,6],
                                 measure[,7]),
                 "method" = c(rep("empirical", dim(measure)[1]),
                              rep("corshrink", dim(measure)[1]),
                              rep("corshrink2-sparse", dim(measure)[1]),
                              rep("corshrink2-loss-0.00001", dim(measure)[1]),
                              rep("corshrink2-loss-0.01", dim(measure)[1]),
                              rep("corshrink2-loss-1", dim(measure)[1]),
                              rep("corshrink2-loss-10", dim(measure)[1]))
)

library(ggplot2)
p <- ggplot(df1, aes(method, frobenius, color = method)) + ylab("Frobenius Distance")
p3 <- p + geom_boxplot() + theme_bw() + theme(legend.position="none") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
p3

```

